---
title: "chapter6"
author: "Julia ROQUIGNY"
date: "2023-12-08"
output: html_document
---

```{r configur, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Part 1: RATS data 

```{r ch6 1}
#original data 
RATS <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/rats.txt", header = TRUE, sep = '\t')

# Read the RATS data
RATSL<-read.csv("D:/OneDrive/Documents/IODS-project/data/RATSL.csv")

# Look at the (column) names of RATS
names(RATSL)

# Look at the structure of BPRS
str(RATSL)

# Print out summaries of the variables

summary(RATSL)

```

install.packages("ggplot")


```{r ch6 2}
library(dplyr)
library(tidyr)

# Factor treatment & subject
RATSL$Group <- factor(RATSL$Group)
RATSL$ID <- factor(RATSL$ID)



#Access the package ggplot2
library(ggplot2)

# Draw the plot
ggplot(RATSL, aes(x =Time, y =Weight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(RATSL$Weight), max(RATSL$Weight)))


```
We can see that the group 2 has an outlier showing high weight value

```{r ch6 3}
# Standardise the variable Weight
RATSL <- RATSL %>%
  group_by(Time) %>%
  mutate(stdWeight =Weight) %>%
  ungroup()

#plot with the standardised Weight
ggplot(RATSL, aes(x =Time, y =stdWeight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  scale_y_continuous(name = "standardized bprs")


```

```{r ch6 4}
# Number of subjects (per group):
n <- 16

library(dplyr)
library(tidyr)
# Summary data with mean and standard error of bprs by treatment and week 
RATSS <- RATSL %>%
  group_by(Group, Time) %>%
  summarise( mean = mean(Weight), se = sd(Weight )) %>%
  ungroup()

# Glimpse the data
glimpse(RATSS)

# Plot the mean profiles
library(ggplot2)
ggplot(RATSS, aes(x = Time, y = mean, linetype = Group, shape = Group)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2,3)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2,3)) +
  #geom_errorbar(aes(ymin=mean-se, ymax=mean+se, linetype="1"), width=0.3) +
  theme(legend.position = c(0.8,0.8)) +
  scale_y_continuous(name = "mean(Weight) +/- se(Weight)")


```
We can observe that the group 2 and 3 range in a similar level, compared to the group 1 which show a much lower Weight mean. 

```{r ch6 5}
library(dplyr)
library(tidyr)
# Create a summary data by treatment and subject with mean as the summary variable (ignoring baseline week 1)
RATSL8S <- RATSL %>%
  filter(Time > 1) %>%
  group_by(Group, ID) %>%
  summarise( mean=mean(Weight) ) %>%
  ungroup()

# Glimpse the data
glimpse(RATSL8S)

# Draw a boxplot of the mean versus treatment
library(ggplot2)
ggplot(RATSL8S, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(Weight), weeks 1-8")


```


```{r ch6 6}
# Create a new data by filtering the outlier and adjust the ggplot code the draw the plot again with the new data
RATSL8S1 <-  RATSL8S %>% filter(mean < 550)

# Draw a new boxplot of the mean versus treatment
library(ggplot2)
ggplot(RATSL8S1, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(Weight), weeks 1-8")

```
The graphics indicate that there is clear differences between the 3 groups. 

```{r ##7}

# Perform a two-sample t-test but we have 3 levels 
#t.test(mean ~ Group, data = RATSL8S1, var.equal = TRUE)

library(dplyr)
library(tidyr)
summary(aov(mean ~ Group, data = RATSL8S1))

# Add the baseline from the original data as a new variable to the summary data
RATSL8S2 <- RATSL8S %>%
  mutate(baseline = RATS$WD1)


# Fit the linear model with the mean as the response 
fit <- lm(mean~baseline+Group, data = RATSL8S2)

# Compute the analysis of variance table for the fitted model with anova()
anova(fit)

```

The anova test results show a significant difference between at least 2 groups. 

Baseline measurements of the outcome variable in a longitudinal study are often correlated with the chosen summary measure and using such measures in the analysis can often lead to substantial gains in precision when used appropriately as a covariate in an analysis of covariance. We can illustrate the analysis on the data using the WD value corresponding to time zero taken prior to the start of treatment as the baseline covariate.
We see that the baseline WD is strongly related to the WD values taken after treatment has begun, but there is still no evidence of a treatment difference even after conditioning on the baseline value.

###Part 2: BPRS data 


```{r ch6 8}
# read in the BPRS data
BPRSL <- read.csv("D:/OneDrive/Documents/IODS-project/data/BPRSL.csv", header = TRUE)

library(dplyr)
# Factor variables treatment and subject
BPRSL$treatment <- factor(BPRSL$treatment)
BPRSL$subject <- factor(BPRSL$subject)

str(BPRSL)
glimpse(BPRSL)

```

```{r ch6 9}
# Draw the plot
ggplot(BPRSL, aes(x =week, y =bprs, linetype = subject,col=subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))
```

```{r ch6 10}
#create a regression model
BPRS_reg<-lm(bprs ~ week + treatment,data=BPRSL)

summary(BPRS_reg)
```

install.packages("Matrix")

install.packages("lme4")


```{r ##11}
library(lme4)
library(Matrix)

#create a random intercept model 
BPRS_ref <- lmer(bprs ~ week + treatment + (1 | subject), data = BPRSL, REML = FALSE)

#summary model 
summary(BPRS_ref)
```
Considering  the subject as a random effect, the model explains a  small amount of the overall standard deviation in the model.

```{r ch6 12}

# create a random intercept and random slope model
BPRS_ref1 <- lmer(bprs ~ week + treatment + (week | subject), data = BPRSL, REML = FALSE)

# print a summary of the model
summary(BPRS_ref1)

# perform an ANOVA test on the two models
anova(BPRS_ref1, BPRS_ref)

```
The  p value of 0.026 indicates that the model BPRS_ref1 is significantly better than the model with subject as a random effect. 

```{r ch6 13}
library(ggplot2)

# create a random intercept and random slope model with the interaction
BPRS_ref2 <- lmer(bprs ~ week + treatment + (week | subject) + week * treatment, data = BPRSL, REML = FALSE)

# print a summary of the model
summary(BPRS_ref2)

# perform an ANOVA test on the two models
anova(BPRS_ref2, BPRS_ref1)

# add Treatment 1/2 to the beginning of each row in the subject column 
BPRSL <- BPRSL %>%
  mutate(subject = ifelse(treatment == 1, paste0("Treatment1", subject), paste0("Treatment2", subject)))

# plot with colors showing the treatment groups
ggplot(BPRSL, aes(x = week, y = bprs, group = subject, col = treatment)) +
  geom_line(aes(linetype = treatment)) +
  scale_color_manual(values = c("darkgreen", "darkblue"))+
  scale_x_continuous(name = "Week", breaks = seq(0, 8, 1)) +
  scale_y_continuous(name = "BPRS") +
  theme(legend.position = "top")

# Create a vector of the fitted values
Fitted <- fitted(BPRS_ref2)

```
```{r ch6 14}
# create a vector of the fitted values
Fitted <- fitted(BPRS_ref2)

# create a new column fitted 
BPRSL <- BPRSL %>% mutate(Fitted = Fitted)

# draw the plot of BPRSL with the Fitted values of bprs
library(ggplot2)
ggplot(BPRSL, aes(x = week, y = Fitted, group = subject)) +
  geom_line(aes(linetype = treatment)) +
  scale_x_continuous(name = "Time", breaks = seq(0, 60, 20)) +
  scale_y_continuous(name = "Fitted bprs") +
  theme(legend.position = "top")


```


