---
title: "Chapter 5"
author: "Julia ROQUIGNY"
date: "2023-12-04"
output: html_document
---

```{r config, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r #1}
#1 Move the country names to rownames (see Exercise 5.5). Show a graphical overview of the data and show summaries of the variables in the data. Describe and interpret the outputs, commenting on the distributions of the variables and the relationships between them

human<-read.csv("D:/OneDrive/Documents/IODS-project/data/create_human.csv")

# remove the column X
human <- dplyr::select(human, -X)

# move the country names to rownames
library(tibble)
human<- column_to_rownames(human, "Country")

# Access GGally
library(GGally)

# visualize the 'human' variables
ggpairs(human,progress=FALSE)

# summary of variables and structure of data
summary(human)
str(human)

```
The data consists of 155 observations with 9 variable. The variables do not look normally distributed except Edu2.FM
We observe strong correlation between: 
- Edu.exp and Life.exp, Mat.Mor (negative correlation) and Ado.Birth (negative correlation)

It shows that countries with long years of schooling , the Life expectancy at birth, the maternal mortality ratio and the Adolescent birth rate are lower



```{r #2}
#2 Perform principal component analysis (PCA) on the raw (non-standardized) human data. Show the variability captured by the principal components. Draw a biplot displaying the observations by the first two principal components (PC1 coordinate in x-axis, PC2 coordinate in y-axis), along with arrows representing the original variables. 

# perform principal component analysis (with the SVD method)
pca_human <- prcomp(human)

# create and print out a summary of pca_human
s <- summary(pca_human)
print(pca_human)

# rounded percentanges of variance captured by each PC
pca_pr <- round(1*s$importance[2, ], digits = 5)

# print out the percentages of variance
print(pca_pr)

# create object pc_lab to be used as axis labels
pc_lab<-paste0(names(pca_pr), " (", pca_pr, "%)")

# draw a biplot
biplot(pca_human, cex = c(0.8, 1), col = c("grey40", "deeppink2"), xlab = pc_lab[1], ylab =  pc_lab[2])

```
```{r #3}
#3 Standardize the variables in the human data and repeat the above analysis. Interpret the results of both analysis (with and without standardizing). Are the results different? Why or why not? Include captions (brief descriptions) in your plots where you describe the results by using not just your variable names, but the actual phenomena they relate to

# standardize the variables
human_std <- scale(human)

# print out summaries of the standardized variables
summary(human_std)

# perform principal component analysis (with the SVD method)
pca_human_std <- prcomp(human_std)

# create and print out a summary of pca_human
s <- summary(pca_human_std)
print(pca_human_std)

# rounded percentanges of variance captured by each PC
pca_pr_std <- round(1*s$importance[2, ], digits = 5)

# print out the percentages of variance
print(pca_pr_std)

# create object pc_lab to be used as axis labels
pc_lab_std<-paste0(names(pca_pr_std), " (", pca_pr_std, "%)")

# draw a biplot
biplot(pca_human_std, cex = c(0.8, 1), col = c("grey40", "deeppink2"), xlab = pc_lab_std[1], ylab =  pc_lab_std[2])

```
 

The results from raw data in comparison to the standardized data look diffrent from each other . In the raw data, GNI takes almost all of the variation.
In the standardized data, the variation is shared between multiple components.We can observe 3 separate groups : 
- Labo.FM and Parli.F
- Ado.Birth and Mat.Mor 
- Edu.Exp, Life.exp, Edu2.FM, GNI
High incomes and education negatively negatively correlate with maternal mortality ratio and the Adolescent birth rate whereas they positively correlate with Life expectancy at birth. 
The % male and females in the labor force and in representatives in parliament  seem to not be associated with  incomes and education. 


```{r #4}
#5 Load the tea dataset and convert its character variables to factors

tea <- read.csv("https://raw.githubusercontent.com/KimmoVehkalahti/Helsinki-Open-Data-Science/master/datasets/tea.csv", stringsAsFactors = TRUE)
library(dplyr)
library(tidyr)

str(tea)
head(tea)


# column names to keep in the dataset

tea_time <-dplyr::select(tea, Tea, How, how, sugar, where, lunch)

# look at the summaries and structure of the data
str(tea_time)

# visualize the dataset
library(ggplot2)
pivot_longer(tea_time, cols = everything()) %>% 
  ggplot(aes(value)) + geom_bar()+facet_wrap("name", scales = "free")+theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))


```

```{r #5}
#Use Multiple Correspondence Analysis (MCA) on the tea data (or on just certain columns of the data, it is up to you!). Interpret the results of the MCA and draw at least the variable biplot of the analysis. You can also explore other plotting options for MCA. Comment on the output of the plots


library(FactoMineR)
mca <- MCA(tea_time, graph = FALSE)

summary(mca)

# visualize MCA
plot(mca, invisible=c("ind"), habillage = "quali",graph.type = "classic")
plot(mca, invisible=c("ind"), habillage = "quali",graph.type="ggplot")
```
 It seems that the type of tea (tea back and/or unpackaged tea) is related to the place where people buy tea.
 For example, people that buy unpackaged tea would buy tea in tea shop. People that buy tea in chain store and tea shop,would  buy both tea bag and unpackaged tea. And people that buy tea bag would buy tea in chain store. 
